---
title: "Stats Final"
output:
  word_document: default
  html_document: default
date: "2024-12-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reading in the data
```{r}
df <- read.csv("allegheny_county_master_file.csv")
```

Creating our sample
```{r}
set.seed(12345)
# using only Coraopolis
df_sub <- df[df$MUNIDESC == "Coraopolis  ",]
# only single family homes
df_sub <- df_sub[df_sub$USEDESC == "SINGLE FAMILY",]
# removing NA values since there aren't many
df_sub <- na.omit(df_sub)

# selecting what columns we want
df_sub <- df_sub[,c("NEIGHCODE","LOTAREA","HOMESTEADFLAG","FAIRMARKETTOTAL","COUNTYTOTAL","LOCALTOTAL","STYLEDESC","STORIES","YEARBLT","EXTFINISH_DESC","ROOFDESC","BASEMENTDESC","GRADEDESC","CDUDESC","TOTALROOMS","BEDROOMS","FULLBATHS","HALFBATHS","HEATINGCOOLINGDESC","FIREPLACES", "BSMTGARAGE", "FINISHEDLIVINGAREA")]

# creating a random sample of 200 observations
ind <- sample(c(1:nrow(df_sub)), size = 200)
sample_data <- df_sub[ind,]
```

Checking sample representation



summary(sample_data)
print("-----------------------------------------------")
summary(df_sub[,c("NEIGHCODE","LOTAREA","HOMESTEADFLAG","FAIRMARKETTOTAL","COUNTYTOTAL","LOCALTOTAL","STYLEDESC","STORIES","YEARBLT","EXTFINISH_DESC","ROOFDESC","BASEMENTDESC","GRADEDESC","CDUDESC","TOTALROOMS","BEDROOMS","FULLBATHS","HALFBATHS","HEATINGCOOLINGDESC","FIREPLACES", "BSMTGARAGE", "FINISHEDLIVINGAREA")])



Checking sample representation for categorical variables
```{r}
# homestead flag *
prop.table(table(sample_data$HOMESTEADFLAG))
prop.table(table(df_sub$HOMESTEADFLAG))
# style desc - make a flag for old style * 
prop.table(table(sample_data$STYLEDESC))
prop.table(table(df_sub$STYLEDESC))
# exterior finish desc * (frame brick other if needed)
prop.table(table(sample_data$EXTFINISH_DESC))
prop.table(table(df_sub$EXTFINISH_DESC))
# neighbor code *
prop.table(table(sample_data$NEIGHCODE))
prop.table(table(df_sub$NEIGHCODE))
# roof desc - shingles flag *
prop.table(table(sample_data$ROOFDESC))
prop.table(table(df_sub$ROOFDESC))
# basement desc - probably don't include in model *
prop.table(table(sample_data$BASEMENTDESC))
prop.table(table(df_sub$BASEMENTDESC))
# grade desc *
prop.table(table(sample_data$GRADEDESC))
prop.table(table(df_sub$GRADEDESC))
# CDU desc - groups based on order *
prop.table(table(sample_data$CDUDESC))
prop.table(table(df_sub$CDUDESC))
# heating and cooling desc *
prop.table(table(sample_data$HEATINGCOOLINGDESC))
prop.table(table(df_sub$HEATINGCOOLINGDESC))
```

Looking for Multicolinearity
```{r}
cor_matrix <- cor(sample_data[,c("LOTAREA","FAIRMARKETTOTAL","LOCALTOTAL","STORIES","YEARBLT","BEDROOMS","FULLBATHS","HALFBATHS","FIREPLACES", "BSMTGARAGE", "FINISHEDLIVINGAREA")])

cor_matrix[((cor_matrix >= .6) | (cor_matrix <= -.6)) & (cor_matrix < 1)]

cor_matrix

# take out total rooms and county total
```

Making Dummy Variables
```{r}
# homestead flag
sample_data$HOMESTEADFLAG <- ifelse(sample_data$HOMESTEADFLAG == "HOM", 1, 0)

# style flag
sample_data$STYLEDESC_OLD <- ifelse(sample_data$STYLEDESC == "OLD STYLE", 1, 0)

# roof flag
sample_data$ROOFDESC_SHINGLE <- ifelse(sample_data$ROOFDESC == "SHINGLE", 1, 0)

# Grade description
sample_data$NEWGRADEDESC <- "Hold"
sample_data[sample_data$GRADEDESC == "AVERAGE -", ]$NEWGRADEDESC <- "AVERAGE"
sample_data[sample_data$GRADEDESC == "AVERAGE +", ]$NEWGRADEDESC <- "AVERAGE"
sample_data[sample_data$GRADEDESC == "AVERAGE", ]$NEWGRADEDESC <- "AVERAGE"
sample_data[sample_data$GRADEDESC == "BELOW AVERAGE", ]$NEWGRADEDESC <- "BELOW"
sample_data[sample_data$GRADEDESC == "BELOW AVERAGE -", ]$NEWGRADEDESC <- "BELOW"
sample_data[sample_data$GRADEDESC == "BELOW AVERAGE +", ]$NEWGRADEDESC <- "BELOW"
sample_data[sample_data$GRADEDESC == "GOOD -", ]$NEWGRADEDESC <- "ABOVE"
sample_data[sample_data$GRADEDESC == "GOOD +", ]$NEWGRADEDESC <- "ABOVE"
```

New Variable Age
```{r}
# how old is the house
sample_data$AGE <- 2024 - sample_data$YEARBLT
```

Transforming for some outliers
```{r}
# lot area
hist(sample_data$LOTAREA, main = "Distribution of Lot Area", ylab = "Lot Area", col = "lightseagreen")
sample_data$LOTAREA_LOG <- log(sample_data$LOTAREA)
hist(sample_data$LOTAREA_LOG, main = "Log(Lot Area)", ylab = "Log(Lot Area)", col = "lightseagreen")

```


```{r}
#sample_data[,c("NEIGHCODE","LOTAREA","HOMESTEADFLAG","FAIRMARKETTOTAL","STYLEDESC_OLD","STORIES","AGE","EXTFINISH_DESC","ROOFDESC_SHINGLE","GRADEDESC","CDUDESC","BEDROOMS","FULLBATHS","HALFBATHS","HEATINGCOOLINGDESC","FIREPLACES", "BSMTGARAGE", "FINISHEDLIVINGAREA")]


model_data <- sample_data[,c("NEIGHCODE","LOTAREA_LOG","HOMESTEADFLAG","FAIRMARKETTOTAL","STYLEDESC_OLD","STORIES","AGE","EXTFINISH_DESC","ROOFDESC_SHINGLE","NEWGRADEDESC","CDUDESC","BEDROOMS","FULLBATHS","HALFBATHS","HEATINGCOOLINGDESC","FIREPLACES", "BSMTGARAGE", "FINISHEDLIVINGAREA")]

```

----------BASELINE MODEL----------
Full Model
```{r}
# fair market total and local total have correlation 1
full_model <- lm(FAIRMARKETTOTAL~., data = model_data)
summary(full_model)

```

QQ plot
```{r}
plot(full_model, which = 2, col = "lightseagreen")
```

KS Test
```{r}
residuals <- full_model$residuals
ks.test(residuals/sd(residuals), "pnorm")
```

Need to try transform target
```{r}
hist(model_data$FAIRMARKETTOTAL, col = "lightseagreen", main = "Distribution of FAIRMARKETTOTAL",
     xlab = "FAIRMARKETTOTAL")
hist(log(model_data$FAIRMARKETTOTAL), col = "lightseagreen", main = "Distribution of log(FAIRMARKETTOTAL)",
     xlab = "log(FAIRMARKETTOTAL)")

```

----------TRANSFORMED MDOEL----------
Full Model - transformed
```{r}
# fair market total and local total have correlation 1
transformed_model <- lm(log(FAIRMARKETTOTAL)~., data = model_data)
summary(transformed_model)
```

QQ plot
```{r}
plot(transformed_model, which = 2, col = "lightseagreen")
```

KS Test
```{r}
residuals <- transformed_model$residuals
ks.test(residuals/sd(residuals), "pnorm")
```

----------BEST SUBSET SELECTION----------
Best subset selection to select the best model
```{r}
#install.packages("leaps")
library(leaps)

best_subset <- regsubsets(log(FAIRMARKETTOTAL) ~., data = model_data, 
                          nvmax = length(coef(transformed_model)-1), method = "exhaustive")

subset_summary <- summary(best_subset)

# Scree plots
plot(subset_summary$adjr2, ylab = "Adjusted R2", xlab = "Number of Variables in Model",
     main = "Adjusted R2 For N Variable Model", col = "lightseagreen", pch = 19)
abline(v = 7, lwd = 2, lty = 2, col = "grey80")
plot(subset_summary$bic, ylab = "BIC", xlab = "Number of Variables in Model",
     main = "BIC For N Variable Model", col = "lightseagreen", pch = 19)
abline(v = 7, lwd = 2, lty = 2, col = "grey80")
plot(subset_summary$cp, ylab = "Cp", xlab = "Number of Variables in Model",
     main = "Cp For N Variable Model", col = "lightseagreen", pch = 19)
abline(v = 7, lwd = 2, lty = 2, col = "grey80")


```

----------FINAL MODEL----------
```{r}
# 7 variable model
coefficients <- coef(best_subset, id = 7)

final_mod <- lm(log(FAIRMARKETTOTAL) ~ LOTAREA_LOG + HEATINGCOOLINGDESC + HOMESTEADFLAG + NEIGHCODE + CDUDESC + FINISHEDLIVINGAREA, data = model_data)

summary(final_mod)

# coefficients
round(coef(final_mod),4)
round(exp(coef(final_mod)),4)
```



Residual Plot - looks evenly distributed and constant variance
```{r}
plot(final_mod, which = 1, col = "lightseagreen")
```

QQ plot - looks roughly normal
```{r}
plot(final_mod, which = 2, col = "lightseagreen")
```

KS Test - Can assume normality
```{r}
residuals_1 <- final_mod$residuals
ks.test(residuals_1/sd(residuals_1), "pnorm")
```












































